Design of a urban sound scenes corpuses with controlled acoustic properties

### 2017-04-24 articleBoston.tex
fix intro

The urban sound environment is the subject of many researches both on perspective \cite{jin_yong_soundwalk_2013} \cite{botteldooren_understanding_2011} and on physical aspects \cite{can_describing_2015} \cite{raimbault_ambient_2003}. These studies can be based on real sound environments either by acoustic listening through a \textit{soundwalk} \cite{adams_soundwalking_2008} or with recording made in city \cite{botteldooren_temporal_2006} or in a laboratory by using sound mixtures resulting from a simulation process \cite{lafay_new_2014}. In the first case, if the urban sound environment have an undeniable ecology validity, it doesn't make it possible to have a repeatable experience nor to propose a controlled framework where the presence and the level of the different sound sources can be adjusted. Then the use of a simulation tool to create specific urban sound environments is a the only way to control the presence of each class. It allows for instance to create urban sound environments by adjusting some parameters \cite{bruce_development_2009} or to test some classification or detection tools \cite{giannoulis_detection_2013} .

The main question is then to get scenes that are realistic enough to be similar to the recordings. Various methods have been proposed. The first approach would be to simulate completely a neighborhood taking into account the architecture of the buildings, the sound dynamic of the different sources present (car, voice, bird, bell \dots) and their propagation to a receptor, as proposed by \cite{cstb_simulation_2015}. If the first renderings are interesting, this method stays complicated to implement with a final rendering that is again too artificial.
The second approach composes sound mixtures from isolated sound associated together. The idea is to consider the urban sound environment as the sum of acoustic events, i.e events with a sufficient sound level to be discernible, superposed to a sound background, i.e a sound constant on the sound mixture whose properties vary slowly during the scene. The difficulty is to create a representative database of isolated sounds with a sufficient quality to not deteriorate the sound quality and to put the acoustic events wisely to compose correct sound environments. The method proposed in \cite{misra_musical_2007} enables to resolve the first issue by extracting the acoustical event directly in real recordings and by manipulating them to re-use them in new mixtures. If the tool has a sufficient number of parameters to control the sound event, their method is limited by the extraction phase: the overlapping between the acoustics events deteriorates the sounds and then the final rendering. In more, the manipulation (duration, frequency range) can create artifacts that can be a limitation for a realistic use. In a similar way, the tool proposed by Davis and Bruce \cite{bruce_development_2009} enables to compose sound mixtures with an isolated sound database composed of recordings made specially for their application. The creation of the urban sound mixtures is made possible with the perceptual evaluation of a panel of listeners of the different sound environments. This method enables to determine which sounds are the most representative or the most linked to an urban sound environment from a perspective point of view.

However, as the objective is to get simulated scenes that are as realistic than recordings, this method can not be used here because if this method proposes a perspective validity it can not provide an ecological one. Thus, this study proposes to create urban sound mixtures from the listening of urban audio recordings and the estimation of some parameters extracted from them to use it in an audio simulation tool. The realism of the composed scene is then tested in a perspective test. The first part focuses on the study of a recordings made in Paris, then the second part deals with the simulation tool and the last part summarizes the result obtains in the perspective test.

\section{Study of realistic scenes}
Real urban noise recordings are listening in details to determine the composition of typical urban sound environment in terms of sound sources, presence and level. The recordings used have been made on the \nth{13} district of Paris (France) in frame of the GRAFIC project context \cite{aumond_sound_2016} during a soundwalk which was designed to cover different types of urban sound environments. The walk was 2.1 km long, consisting of 19 stop points of 3 to 5 minutes of listening each (see figure \ref{fig:soundwalk}). It was roamed on two days (03/23/2015 and 03/30/2015) twice a day (on the morning and on the afternoon) and in one direction (from West to East, WE) and then in the other (from East to West, EW). The acquisition system used was a ASA SENS and was carry on the backpack of the operator which was walking among the participants. In the end, 74 audio in a wav format and sampled at 44,1 kHz are available and serve as a base for the study. More details can be found on  \cite{aumond_modelling_2017}. \\

@@

\section{Creation of realistic simulated urban scenes}
\subsection{Presentation of the web-simulator \textit{simScene}}
\textit{simScene} \cite{rossignol_simscene:_2015} is a web-simulator that creates sound mixtures in a wav format by superposing audio which comes from an isolated sound database composed of two categories: the brief sounds (from 1 to 20 seconds)
that can be salient are belong to the \textit{event} category and all the sound that are long and whose acoustic properties do not vary are put in the \textit{background} category. In each category, the sounds are then regrouping in some sound classes (bird, car, foot steps \dots). The software enables to control for each sound classes some parameters as the number of events of each class that appears in the mixture, the elapsed time between each sample of a same class or the existence of a fade in and a fade out \dots. Each parameters are completed with a standard deviation bringing some random behavior between the scenes. \\

@@

\section{Perceptual test}
\subsection{Creation of the test}
A perspective test is created where a panel of listeners evaluates the realism of a mix of real and replicated scenes on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}). Since it is not possible for each panelist to evaluate all the real and replicated scenes, in order to limit the duration of the test and to keep the judge's concentration constant, each listens a panel of 20 sound scenes composed of 10 real sounds scenes and 10 replicated. Furthermore, for the auditory comfort of the panelist, all the scenes are normalized to the same sound level, chosen at 65 dB.\\

@@

\section*{Acknowledge}
We would like to thank Pierre Aumond and Catherine Lavandier from the University of Cergy-Pontoise for transmitting us the data of the \textit{Grafic} project.

>>>>

The urban sound environment is studied both from the perception \cite{jin_yong_soundwalk_2013} \cite{botteldooren_understanding_2011} and from the physical point of view \cite{can_describing_2015} \cite{raimbault_ambient_2003}. These studies can be based on real sound environments either by acoustic listening through a \textit{soundwalk} \cite{adams_soundwalking_2008} or with recording made in city \cite{botteldooren_temporal_2006} or in a laboratory by using sound mixtures resulting from a simulation process \cite{lafay_new_2014}. In the first case, if the urban sound environment have an undeniable ecology validity, it doesn't make it possible to have a repeatable experience nor to propose a controlled framework where the presence and the level of the different sound sources can be adjusted. Then the use of a simulation tool to create specific urban sound environments is a useful way to control the degree of contribution of each class of acoustical sources. It allows us to create urban sound environments by adjusting some parameters \cite{bruce_development_2009} or to evaluate the performance of some classification or detection tools \cite{giannoulis_detection_2013}.

A critical question is then to obtain scenes that are realistic enough, thus to be considered as similar to some reference recordings. To do so, various methods have been proposed. The first approach would be to simulate completely a neighborhood taking into account the architecture of the buildings, the sound dynamic of the different sources present (car, voice, bird, bell \dots) and their propagation to a receptor, as proposed by \cite{cstb_simulation_2015}. If the first renderings are interesting, this method remains complex to implement.
A more straitghtforward approach composes sound mixtures from isolated sound mixed together. The idea is to consider the urban sound environment as the sum of acoustic events, i.e. events with a sufficient sound level to be discernible, superposed to a sound background, i.e a sound constant on the sound mixture whose properties vary slowly during the scene. The difficulty here is to have a representative database of isolated sounds with a sufficient quality not to compomise the resulting sound quality and to sequence the acoustic events wisely.

The method proposed in \cite{misra_musical_2007} enables to resolve the first issue by extracting the acoustical event directly in real recordings and by manipulating them to re-use them in new mixtures. If the tool has a sufficient number of parameters to control the sound event, their method is limited by the extraction phase: the overlapping between the acoustics events deteriorates the sounds and then the final rendering. In more, the manipulation (duration, frequency range) can create artifacts that can be a limitation for a realistic use. In a similar way, the tool proposed by Davis and Bruce \cite{bruce_development_2009} enables to compose sound mixtures with an isolated sound database composed of recordings made specially for their application. The creation of the urban sound mixtures is made possible with the perceptual evaluation of a panel of listeners of the different sound environments. This method enables to determine which sounds are the most representative or the most linked to an urban sound environment from a perspective point of view.

However, as our objective here is to get simulated scenes that are as realistic as possible, this method can not be considered because if this method proposes a perspective validity it can not provide an ecological one. Thus, this study proposes to create urban sound mixtures from the listening of urban audio recordings and the estimation of some high level sequencing parameters extracted from the recordings to tune an audio simulation tool. The realism of the simulated scenes is then evaluated using a perspective test.

The remaining of the paper is organized as follows, Section \ref{sec:study} focuses on the study of recordings made in Paris. Section \ref{sec:simulation} deals with the simulation tool and Section \ref{sec:test} summarizes the result obtained in the perspective test.

\section{Study of realistic scenes} \label{sec:study}
Real urban noise recordings are listening in details to determine the composition of typical urban sound environment in terms of sound sources, presence and level. The recordings used have been made on the \nth{13} district of Paris (France) in frame of the GRAFIC project context \cite{aumond_sound_2016} during a soundwalk which was designed to cover different types of urban sound environments. The walk was 2.1 km long, consisting of 19 stop points of 3 to 5 minutes of listening each (see figure \ref{fig:soundwalk}). It was roamed on two days (03/23/2015 and 03/30/2015) twice a day (on the morning and on the afternoon) and in one direction (from West to East, WE) and then in the other (from East to West, EW). The acquisition system used was a ASA SENS and was carry on the backpack of the operator which was walking among the participants. In the end, 74 audio in a wav format and sampled at 44,1 kHz are available and serve as a base for the study. More details can be found on  \cite{aumond_modelling_2017}. \\

@@

\section{Simulation of realistic simulated urban scenes} \label{sec:simulation}
\subsection{Presentation of the web-simulator \textit{simScene}}
\textit{simScene} \cite{rossignol_simscene:_2015} is a web-simulator that creates sound mixtures in a wav format by superposing audio which comes from an isolated sound database composed of two categories: the brief sounds (from 1 to 20 seconds)
that can be salient are belong to the \textit{event} category and all the sound that are long and whose acoustic properties do not vary are put in the \textit{background} category. In each category, the sounds are then regrouping in some sound classes (bird, car, foot steps \dots). The software enables to control for each sound classes some parameters as the number of events of each class that appears in the mixture, the elapsed time between each sample of a same class or the existence of a fade in and a fade out \dots. Each parameters are completed with a standard deviation bringing some random behavior between the scenes. \\

@@

\section{Perceptual test} \label{sec:test}
\subsection{Creation of the test}
A perspective test is created where a panel of listeners evaluates the realism of a mix of real and replicated scenes on a 7-point scale (1 is \textit{not realistic at all}, 7 is \textit{very realistic}). Since it is not possible for each panelist to evaluate all the real and replicated scenes, in order to limit the duration of the test and to keep the judge's concentration constant, each listens a panel of 20 sound scenes composed of 10 real sounds scenes and 10 replicated. Furthermore, for the auditory comfort of the panelist, all the scenes are normalized to the same sound level, chosen at 65 dB.\\

@@

\section*{Acknowledgements}
We would like to thank Pierre Aumond and Catherine Lavandier from the University of Cergy-Pontoise for transmitting us the data of the \textit{Grafic} project.
